---
title: "p8105_hw2_kx2173"
author: "ke"
date: "10/8/2021"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet.

```{r trash_wheel}
trash_wheel = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         sheet = "Mr. Trash Wheel", 
                         range = "A2:N535")%>%
  janitor::clean_names()%>%
  drop_na(dumpster)%>%
  mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019.

```{r combined_precipitation}
precipitation_2018 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         sheet = "2018 Precipitation", 
                         range = "A2:B14")%>%
  janitor::clean_names()%>%
  drop_na(month, total)%>%
  mutate(year = "2018")%>%
  relocate(year)



precipitation_2019 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                         sheet = "2019 Precipitation", 
                         range = "A2:B14")%>%
  janitor::clean_names()%>%
  drop_na(month, total)%>%
  mutate(year = "2019")%>%
  relocate(year)

combined_precipitation = 
  bind_rows(precipitation_2018, precipitation_2019)%>%
  mutate(
    month = as.integer(month),
    month = month.name[month])

class(pull(combined_precipitation, month)) # To check if "month" variable is a character variable.
```

Write a paragraph about data:

In Trash Wheel data, it has key variables: `r names(trash_wheel)`, with number of `r ncol(trash_wheel)` and has `r nrow(trash_wheel)` observations.
In precipitation data for 2018 and 2019, it has variables: `r names(combined_precipitation)`, and total number of observation is `r nrow(combined_precipitation)`.
The total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`.
The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trash_wheel, year == 2019), sports_balls))`.


## Problem 2

First, read and clean data in pols_month.

```{r pols_month}
pols_month = read_csv("./fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(mon, c("year", "month", "day"))%>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.name[month],
    president = ifelse(prez_dem == "1", "dem", "gop")
    )%>%
  select(- day, -prez_dem, -prez_gop)%>%
relocate(year, month, president)
```

Second, clean the data snp.csv as well, arrange according to year and month, and organize columns.

```{r snp}
snp = read_csv("./fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  mutate(date = lubridate::mdy(date))%>%
  separate(date, c("year", "month", "day"))%>%
  mutate(
   month = as.integer(month),
   month = month.name[month],
   year = as.integer(year),
   year = ifelse(year>2021, year - 100, year)
  )%>%
  select(-day)%>%
  arrange(year, month)%>%
  relocate(year,month)
```

Third, read and tidy the unemployment data.

```{r unemployment_tidy}
unemployment_data = read_csv("./fivethirtyeight_datasets/unemployment.csv")%>%
  janitor::clean_names()
colnames(unemployment_data) = c("year","January","February","March","April","May","June","July","August","September","October","November","December")

unemployment_tidy = 
  unemployment_data %>%
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemployment_percentage"
  )%>%
  drop_na()%>%
  relocate(year, month)
```

Now, join the datasets.

```{r}
pols_snp = left_join(pols_month, snp, by = c("year", "month"))

pols_snp_unemp = left_join(pols_snp, unemployment_tidy, by = c("year", "month"))
```

A short paragraph about these data sets.



## Problem 3

load and tidy data.

```{r baby_tidy}
baby_name = read_csv("./baby_names/Popular_Baby_Names.csv")%>%
  janitor::clean_names()%>%
  mutate(
    childs_first_name = tolower(childs_first_name),
    gender = toupper(gender),
    ethnicity = toupper(ethnicity)
    )

unique(pull(baby_name, ethnicity)) # Here we find something needs to be uniform

baby_tidy = baby_name[!duplicated(baby_name), ]%>%
  mutate(
    ethnicity = recode(ethnicity, 
                       `ASIAN AND PACI` = "ASIAN AND PACIFIC ISLANDER",
                       `BLACK NON HISP` = "BLACK NON HISPANIC",
                       `WHITE NON HISP` = "WHITE NON HISPANIC"))
```

Produce a table showing rank in popularity of the name "Olivia" as a female baby name.

```{r olivia_rank}
olivia_rank = baby_tidy%>%
  filter(gender == "FEMALE", childs_first_name == "olivia")%>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )%>%
  select(-gender)%>%
  relocate("childs_first_name")

knitr::kable(olivia_rank, caption = "Rank in popularity of the name Olivia as a female baby name")
```

Produce a similar table showing the most popular name among male children

```{r popular_male_name}
popular_male_name = baby_tidy%>%
  filter(gender == "MALE", rank == "1")%>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )%>%
  select(
    -gender,
    -rank
  )

knitr::kable(popular_male_name, caption = "The most popular name among male children over time")
```

Finally, produced a scatter plot showing the number of children with a name against the rank in popularity of that name.

```{r}
white_male_name = baby_tidy%>%
  filter(gender == "MALE",
         ethnicity == "WHITE NON HISPANIC",
         year_of_birth == "2016")

plot_final = 
  ggplot(white_male_name, aes(x = rank, y = count)) + geom_point() + 
  labs(title = "Number of children with a name against the rank in popularity of that name",
       subtitle = "For male, white non-hispanic children born in 2016",
       x = "The rank in popularity of the certain name",
       y = "Number of children with a name"
       )
```

